Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers

### 2016-06-25 notes/paper/paper.tex
Incremental updates to the paper. Tweaks to the figures and wording on some parts.

If we let $\textbf{a} = \{a_i\}_{i=0}^{N}$ represent our demeaned acceleration data over a specified time window in three dimensions then the covariance matrix of our acceleration data is represented by the real-symmetric $3 \times 3$ matrix $Q = \textbf{a} \textbf{a}^T$.
The resulting unit length eigen vectors $V = (\textbf{v}^1,\textbf{v}^2,\textbf{v}^3)$ with their respective eigen values $\lambda^1 \geq \lambda^2 \geq \lambda^3$ of this matrix are orthonormal and point in the directions where the data varies most in descending order for each orthogonal direction.
% \ref{bishop_2006}
\begin{equation} \label{project_eq}
   \textbf{a}' = V^T \textbf{a}
\end{equation}
%
We now represent our acceleration data as $\textbf{a}'$ after projecting our acceleration data $\textbf{a}$ onto our rotated axis $V$ by \eqref{project_eq}. We make the assumption that our data is most dominate in two dimensions i.e. $\lambda^3 << \lambda^2$ so it is safe to assume that
%
\begin{equation} \label{magnitude_eq}
||(a^1,a^2,a^3)|| = ||(a'^1,a'^2,a'^3)|| \sim = (a'^1)^2 + (a'^2)^2
\end{equation}
$ \forall a \in \textbf{a}. $ \\
%
Equality comes from the orthonormality of $V$. The margin of error in the approximation is not important since we are only interested in processing in the two dimensions.
By now we have preserved as much information that is contained in our acceleration data as possible while also reducing our dimension from three to two.
By projecting our data onto the plain defined by the two most dominant eigen vectors, we effectively reduce our representation of our data to two dimensions.

@@

For our application, we use a triangle filter prior to downsampling. This is achieved by convolving our data with the following weights:
%
\begin{equation} \label{triangle_filter_weights}
w_k^N =
\begin{cases}
  \frac{k+1}{N^2}, & \text{if}\ 0 \leq k < N \\
  \frac{2N - (k + 1)}{N^2}, & \text{if} N \leq k < 2N - 1
\end{cases}
\end{equation}
%
For a decimation value of $N$. Then the data is downsampled by $N$. This results in:
%
\begin{equation} \label{decimated_signal}
s_k^1 = \sum_{p=0}^{2N-1} w_{p+N(1-k)}^N s_{kN - (p+1)}^0 ?
\end{equation}
%
where $\{s_k^1\}_{k=0}^{M/N}$? is the decimated signal of $\{s_k^0\}_{k=0}^M$?.
\subsection{Training the Module}
We are able to identify an individual\textquotesingle s activity by matching the signal to a reference template.
For our application, we develop a reference signature for our matched filter directly from the person\textquotesingle s motion signature.
We do this by having the individual perform the specific activity over a specified amount of time and select the sub-interval of time that we find best represents the user's motion signature.
For a given activity, we attempt to find the representative periodic signal $s$ of our individual’s activity to be our template for our matched-filter.

@@

We chose this method for its simplicity and robustness against outliers in our training set. We anticipate that the user will have idle motion at the start and end of their training duration when they activate the training mode and with possible disruptions during the process.
%
\section{Results}
%
\subsection{Algorithm Performance}
%
The Kinetisense sensor was used for the performance analysis in this paper.
This unit has a three-axis accelerometer with an acceleration range of -5g to +5g and a sampling rate of 128Hz.
Samples of walking, jogging, and bicycling have been recorded with this unit attached to the ankle of the user for the purpose of analyzing the performance of the algorithms described in this paper.
A separate instance of walking data has been used specifically for training the classifier.
In figure \ref{fig:MF_performance}, \subref{fig:frac_walking_jogging} and \subref{fig:frac_walking_bicycling} show the fraction of the correlation results that exceed the given correlation threshold when comparing the walking data reference to the given activity.
As we would expect, walking data matches better to walking data than the other activities.
ROC curves were produced from these corresponding results in \subref{fig:roc_walking_jogging} and \subref{fig:roc_walking_bicycling} to illustrate the effectiveness of the matched-filter as a classifier.
Figure \ref{fig:MF_performance} shows the comparison to bicycling.
Because the bicycling motions are considerably different from walking relative to jogging, we would expect better performance from bicycling.
%
\begin{figure}[!ht]
  %\captionsetup[subfigure]{labelformat=empty}
  \centering
  \subfloat[\label{fig:frac_walking_jogging}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_jogging.png}}
  \quad
  \subfloat[\label{fig:frac_walking_bicycling}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_bicycling.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_jogging}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_run.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_bicycling}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_bike.png}}
  \caption{Matched-filter performance comparing jogging and bicycling data against a reference for walking. Results produced with decimation 8}
  \label{fig:MF_performance}
\end{figure}
%
%\FloatBarrier
reference to figure \ref{fig:MF_performance}
%
\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies on the MSP432 micro-controller processor comparing the timing between using one vs two references for classification with the matched-filter. The time window used for activity classification is 4.0 seconds}
   \label{fig:timing}
\end{figure}
%
\subsection{Throughput Performance}
%
To assess the real-time processing potential of the algorithms outlined in this paper, a throughput assessment was performed on the MSP432P401R microcontroller with varying parameters.
The Texas Instrument - MSP432P401R is a low power plus performance microcontroller that is based on the ARM 32-Bit Cortex-M4F processor.
Some of the notable features of the MSP432P410R is having a clock frequency of 48 MHz, 64 kB SRAM, 256 kB Flash Memory, and has an IEEE 754-compliant single-precision Floating-Point Unit (FPU) and Memory Protection Unit (MPU).
The single-precision processing capability on this unit provides us with a sufficiently large dynamic range that fixed-point processing cannot provide.
The 64 kB of SRAM provided sufficient space to process the results found in figure \ref{fig:timing} and
the 256 kB is more than enough flash memory to store the program and data references.
The clock frequency operates at 48 MHz which allows the microcontroller to have a high performance by allowing a high number of instruction cycles that ultimately allows for faster code execution.

@@

Randomized data was used for the purpose of analysing the throughput performance on the device while varying the parameters.
Figure \ref{fig:multiple_comparison} shows execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process while operating at decimation $2$.
Because the FFT dominates the processing load, and was optimized for even powered factors, one can observe better throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even valued time lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than the $50$Hz.
Figure \ref{fig:timing_refs} shows execution time over a $4.0$ second window at decimation $8$ while varying the sampling frequency and number of references. From this, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\%.
%
\section{Conclusion}
conclusion text.
%
\appendices
%
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi
%
\bibliographystyle{plain}
\bibliography{reference}
\end{document}

>>>>

If we let $\textbf{a} = \{a_i\}_{i=0}^{N}$ represent our demeaned acceleration data over a specified time window in three dimensions then the covariance matrix of our acceleration data is represented by the real-symmetric $3 \times 3$ matrix $Q = \textbf{a} \textbf{a}^T$.
The resulting unit length column eigenvectors of $Q$ are $V = (\textbf{v}^1,\textbf{v}^2,\textbf{v}^3)$ with their respective eigenvalues $\lambda^1 \geq \lambda^2 \geq \lambda^3$ of this matrix are orthonormal and point in the directions where the data varies most in descending order for each orthogonal direction.
% \ref{bishop_2006}
\begin{equation} \label{project_eq}
   \textbf{a}' = V^T \textbf{a}
\end{equation}
%
We now represent our acceleration data as $\textbf{a}'$ after projecting our acceleration data $\textbf{a}$ onto our rotated axis $V$ by \eqref{project_eq}. If the data is mostly dominate in the two primary dimensions, i.e. $\lambda^3 << \lambda^2$, then it is safe to assume that
%
\begin{equation} \label{magnitude_eq}
||(a^1,a^2,a^3)||^2 = ||(a'^1,a'^2,a'^3)||^2 \sim = (a'^1)^2 + (a'^2)^2
\end{equation}
$ \forall a \in \textbf{a}. $ \\
%
Equality comes from the orthonormality of $V$. The margin of error in the approximation is not important since we are only interested in processing in the two dimensions.
By now we have preserved as much information that is contained in our acceleration data as possible while also reducing our dimension from three to two.
By projecting our data onto the plain defined by the two most dominant eigen vectors, we effectively reduce our representation of our data to two dimensions.

@@

For our application, we use a triangle filter prior to downsampling. This is achieved by convolving our data with the following weights:
%
\begin{equation} \label{triangle_filter_weights}
w_k^D =
\begin{cases}
  \frac{k+1}{D^2}, & \text{if }\ 0 \leq k < D \\
  \frac{2D - (k + 1)}{D^2}, & \text{if } D \leq k < 2D - 1 \\
  0, & \text{otherwise}
\end{cases}
\end{equation}
%
For a decimation value of $D$. Then the data is downsampled by $D$. This results in:
%
\begin{equation} \label{decimated_signal}
s_k^1 = \sum_{p=0}^{2N-1} w_{p+N(1-k)}^N s_{kN - (p+1)}^0 ?
\end{equation}
%
where $\{s_k^1\}_{k=0}^{M/D}$? is the decimated signal of $\{s_k^0\}_{k=0}^M$?.
\subsection{Training the Module}
We are able to identify an individual\textquotesingle s activity by matching the signal to a reference template.
For our application, we develop a reference signature for our matched filter directly from the person\textquotesingle s motion signature.
We do this by having the individual perform the specific activity over a specified amount of time and select the sub-interval of time that we find best represents the user's motion signature.
For a given activity, we attempt to find the representative periodic signal $s$ of our individual’s activity to be our template for our matched-filter.

@@

We chose this method for its simplicity and robustness against outliers in our training set. We anticipate that the user will have idle motion at the start and end of their training duration when they activate the training mode and with possible disruptions during the process.
%
\section{Results}
%
\subsection{Algorithm Performance}
%
The Kinetisense sensor was used for the algorithm performance analysis in this paper.
This unit has a three-axis accelerometer with an acceleration range of -5g to +5g and a sampling rate of 128Hz.
Samples of walking, jogging, and bicycling have been recorded with this unit attached to the ankle of the user.
A separate instance of walking data has been used specifically for training the classifier.
%
\begin{figure}[!ht]
  %\captionsetup[subfigure]{labelformat=empty}
  \centering
  \subfloat[\label{fig:frac_walking_jogging}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_jogging.png}}
  \quad
  \subfloat[\label{fig:frac_walking_bicycling}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_bicycling.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_jogging}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_run.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_bicycling}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_bike.png}}
  \caption{Matched-filter performance comparing jogging and bicycling data against a reference for walking. Results produced with decimation 8. Figures (c) and (d) are the corresponding ROC curves produced from (a) and (b) respectivley.}
  \label{fig:MF_performance}
\end{figure}
%
Figures \ref{fig:frac_walking_jogging} and \ref{fig:frac_walking_bicycling} show the fraction of the correlation results that exceed the given correlation threshold when comparing the walking data reference to the given activity.
As we would expect, walking data matches better to walking data than the other activities.
ROC curves were produced from these corresponding results in \subref{fig:roc_walking_jogging} and \subref{fig:roc_walking_bicycling} to illustrate the effectiveness of the matched-filter as a classifier.
Figure \ref{fig:MF_performance} shows the comparison to bicycling.
Because the bicycling motions are considerably different from walking relative to jogging, we would expect better performance from bicycling.
%
%\FloatBarrier
%
\subsection{Throughput Performance}
%
To assess the real-time processing potential of the algorithms outlined in this paper, a throughput assessment was performed on the MSP432P401R microcontroller with varying parameters.
The Texas Instrument - MSP432P401R is a low power plus performance microcontroller that is based on the ARM 32-Bit Cortex-M4F processor.
Some of the notable features of the MSP432P410R is having a clock frequency of 48 MHz, 64 kB SRAM, 256 kB Flash Memory, and has an IEEE 754-compliant single-precision Floating-Point Unit (FPU) and Memory Protection Unit (MPU).
The single-precision processing capability on this unit provides us with a sufficiently large dynamic range that fixed-point processing cannot provide.
The 64 kB of SRAM provided sufficient space to process the results found in figure \ref{fig:timing} and
the 256 kB is more than enough flash memory to store the program and data references.
The clock frequency operates at 48 MHz which allows the microcontroller to have a high performance by allowing a high number of instruction cycles that ultimately allows for faster code execution.

@@

%
\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies, time-window lengths, and number of references used for the matched-filter on the MSP432 micro-controller processor. The time window used for (b) is 4.0 seconds.}
   \label{fig:timing}
\end{figure}
%
Randomized data was used for the purpose of analysing the throughput performance on the device while varying the parameters.
Figure \ref{fig:multiple_comparison} shows execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process while operating at decimation $2$.
Because the FFT dominates the processing load, and was optimized for even powered factors, one can observe better throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even valued time lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than the $50$Hz.
Figure \ref{fig:timing_refs} shows execution time over a $4.0$ second window at decimation $8$ while varying the sampling frequency and number of references. From this, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\%.
%
\section{Conclusion}
conclusion text.
%
\appendices
%
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi
%
\bibliographystyle{plain}
\bibliography{reference}
\end{document}
